---
title: "Coastal Health Data Explorer Architecture (For Developers)"
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(CoastalHealth)
```


The Coastal Health Data Explorer is an open source package and the authors welcome any suggestions/developments. In order to support future development, here we provide an overview of the package architecture to help potential developers navigate the package.

## Shiny modules

The app is divided into the following shiny modules, each serving a specific purpose

* mod_area - Management catchment area selection
* mod_layerpopup - Manages the popup table that shows the user selected data layers and provides links for viewing/downloading them
* mod_layerselect - Displays and manages the layer selection tree in the app sidebar
* mod_map - Displays the basemap and any additional layers.
* mod_wbview - Displays toggles in the app sidebar for showing/hiding additional layers.

## Package environment

The passing of internal data between modules, and other internal package varibales are stored in a package enviornment. This environment is created by `env_setup()` and  initally contains placeholders for all internal variables. The `env_setup()` function is also used to reset the package environent to the default state when required.

